### 2021-04-12 paper/ase21/analyzer.tex
[ASE21] temporal dump

In this section, we explain the detail of the type analyzer of $\tool$. We first
formally define a modified $\ires$ and explain how to perform type analysis
for the modified $\ires$.  Besides, we explain a refinement of type analysis
based on conditions of assertions and branches to increase analysis precision.

@@

$\ires$ is an untyped intermediate representation introduced by \citet{jiset}.
To make it more suitable for type analysis, we modify it as a label-based
language as follows:
\[
  \begin{array}{l@{~}r@{~}c@{~}r@{~}r@{~}l}
    \text{Functions}
    &\funcset&\ni&\func&::=&\kwdef \; \x (\x^*, \x^*). \; \lab\\

@@

    \text{Expressions}
    &&&\expr&::=&
    \tname^? \; \kwcl [\x: \expr]^* \kwcr \mid
    \kwsl \expr^* \kwsr \mid
    \expr: \ty \mid
    \refer \kwexists \\

@@

    \text{Types}
    &\tyset&\ni&\ty&::=&\tname \mid \tjs \mid \tprim \mid
    \undefval \mid \nullval \mid \tbool\\

    &&&&\mid&
    \tnumeric \mid \tnum \mid \tbigint \mid \tstr \mid \tsymb \\
  \end{array}
\]
A modified $\ires$ program $\prog = (\funcs, \getlab, \nextlab)$ is a triple of
a set of functions $\funcs \subseteq \and \funcset$ and two finite mappings
$\getlab$ and $nextlab$.  The getter mapping $\getlab: \labset \finmap \instset$
represents instructions attached to labels, and the next mapping $\nextlab:
\labset \finmap \labset$ denotes next labels.  A function $\kwdef \; \f (\x^*,
\y^*). \; \lab \in \funcset$ consists of a name $\f$, normal parameters $\x^*$,
optional parameters $\y^*$, and a body label $\lab$.  A label $\lab \in \labset$
denotes a program point.  An instruction $\inst$ is a variable declaration, a
function call, an assertion, a branch, a return, or an reference update.  Each
invocation of an abstract algorithm in ECMAScript is compiled to a function call
instruction with a newly introduced temporal variable.  We represent loops using
branch instructions with cyclic pointing of labels in the next mapping
$\nextlab$.  A reference $\refer$ is a variable $\x$ or a field access $\refer
\kwsl \expr \kwsr$.  An expression $\expr$ is a record, a list, a type check, an
existence check, a binary operation, a unary operation, a reference, a constant,
or a primitive, which is either $\undefval$, $\nullval$, a Boolean $\bool$, a
Number $\num$, a BigInt $\bigint$, a String $\str$, or a Symbol $\symb$.

@@

A type $\ty \in \tyset$ is a norminal type $\tname$ or a pre-defined type;
$\tjs$ denotes JavaScript values, $\tprim$ primitives, $\tnumeric$ either Number
or BigInt values, and $\tnum$, $\tbigint$, $\tstr$, and $\tsymb$ denote Number,
BigInt, String, and Symbol values, respectively.  A norminal type $\tname$ is a
production name for ASTs or a record type name used in abstract algorithms.  The
subtype relation $\subtype \subseteq \tyset \times \tyset$ between types is
described in Figure~\ref{fig:subtype}; the directed edge from $\ty'$ to $\ty$
denotes $\ty' \subtype \ty$ and the relation is reflexive and transitive.  The
subtype relation is dependent on the syntax and defined record types in
ECMAScript.  We manually model subtypes between record types but automatically
extract subtypes for AST types.  For example, consider the syntax-directed
abstract algorithm at the right bottom in Figure~\ref{fig:example}.  The
non-terminal \textit{BindingElement} is the unique token in a alternative of
\textit{FormalParameter} thus we automatically extract the subtype relation:
\textit{BindingElement} $\subtype$ \textit{FormalParameter}.  Based on this
subtype relation, a type check expression $\expr: \ty$ checks whether the
evaluation result of $\expr$ has a type $\ty'$ satisfying $\ty' \subtype \ty$.
We represent semantics of the modified $\ires$ as a state transition system
$(\stset, \trans, \istset)$ with the transition relation $\trans \subseteq
\stset \times \stset$.  For presentation brevity, we omit its detail in this
paper and include it in a companion report~\inred{\cite{report}}.
% \[
%   \begin{array}{l@{~}r@{~}c@{~}r@{~}r@{~}l}
%     \text{States}
%     &\st&\in&\stset&::=&\ctxtset^* \times \labset \times \envset\\
% 
%     \text{Contexts}
%     &\ctxt$\in$\ctxtset&::=&\labset \times \varset \times \envset\\
% 
%     \text{Environments}
%     &\env$\in$\envset&::=&\varset \finmap \valset\\
% 
%     \text{Values}
%     &\val$\in$\valset&::=& \addrset \uplus \addr \uplus \primset\\\
%   \end{array}
% \]

@@

Abstract Interpretation~\cite{ai1977, ai1992}

\begin{figure*}[t]
  \centering
  \[
    \todo
  \]
  \vspace*{-1em}
  \caption{Abstract .}
  \label{fig:example}
\end{figure*}

>>>>

In this section, we explain the detail of the type analyzer of $\tool$. We first
formally define a modified $\ires$ and explain how to perform type analysis for
the modified $\ires$.  Moreover, we also explain a refinement of type analysis
based on conditions of assertions and branches to increase analysis precision.

@@

$\ires$ is an untyped intermediate representation introduced by \citet{jiset}.
To make it more suitable for type analysis, we modify it as a label-based
language as follows:
\small
\[
  \begin{array}{l@{~}r@{~}c@{~}r@{~}r@{~}l}
    \text{Functions}
    &\funcset&\ni&\func&::=&\kwdef \; \x (\x^*, \x^*). \; \lab\\

@@

    \text{Expressions}
    &&&\expr&::=&
    \tname \; \kwcl [\x: \expr]^* \kwcr \mid
    \clist{\expr^*} \mid
    \expr: \ty \mid
    \refer \kwexists \\

@@

    \text{Types}
    &\tyset&\ni&\ty&::=&\tname \mid \clist{} \mid \clist{\ty} \mid \tjs \mid
    \tprim \mid \undefval \mid \nullval\\

    &&&&\mid&
    \tbool \mid \tnumeric \mid \tnum \mid \tbigint \mid \tstr \mid \tsymb \\
  \end{array}
\]
A modified $\ires$ program $\prog = (\funcs, \getlab, \nextlab)$ is a triple of
a set of functions $\funcs \subseteq \and \funcset$ and two finite mappings
$\getlab$ and $nextlab$.  The getter mapping $\getlab: \labset \finmap \instset$
represents instructions attached to labels, and the next mapping $\nextlab:
\labset \finmap \labset$ denotes next labels.  A function $\kwdef \; \f (\x^*,
\y^*). \; \lab \in \funcset$ consists of a name $\f$, normal parameters $\x^*$,
optional parameters $\y^*$, and a body label $\lab$.  A label $\lab \in \labset$
denotes a program point.  An instruction $\inst$ is a variable declaration, a
function call, an assertion, a branch, a return, or an reference update.  Each
invocation of an abstract algorithm in ECMAScript is compiled to a function call
instruction with a newly introduced temporal variable.  We represent loops using
branch instructions with cyclic pointing of labels in the next mapping
$\nextlab$.  A reference $\refer$ is a variable $\x$ or a field access $\refer
\kwsl \expr \kwsr$.  An expression $\expr$ is a record, a list, a type check, an
existence check, a binary operation, a unary operation, a reference, a constant,
or a primitive, which is either $\undefval$, $\nullval$, a Boolean $\bool$, a
Number $\num$, a BigInt $\bigint$, a String $\str$, or a Symbol $\symb$.

@@

A type $\ty \in \tyset$ is a norminal type $\tname$ or a pre-defined type;
$\tjs$ denotes JavaScript values, $\tprim$ primitives, $\tnumeric$ either Number
or BigInt values, and $\tnum$, $\tbigint$, $\tstr$, and $\tsymb$ denote Number,
BigInt, String, and Symbol values, respectively.  A norminal type $\tname$ is
either 1) a \textit{production type} for ASTs or 2) a \textit{record type} name
used in abstract algorithms, and ECMAScript describes its fields with
pre-defined or possible values.  A production type has corresponding
syntax-directed algorithms as its fields and a record type has specific fields
with possible values described in the corresponding table in ECMAScript.  For
example, the ``Table 10: References Record Fields'' in 
``Reference'' record has

Each norminal type has own fields 
ECMAScript describes fields of each norminal type with their possible values in
table structures ad
scattered tables or syntax-directed algorithms
tables or 
It has pre-defined fields
It has fields

The subtype relation $\subtype \subseteq \tyset \times \tyset$ between types is
described in Figure~\ref{fig:subtype}; the directed edge from $\ty'$ to $\ty$
denotes $\ty' \subtype \ty$ and the relation is reflexive and transitive.  The
subtype relation is dependent on the syntax and defined record types in
ECMAScript.  We manually model subtypes between record types but automatically
extract subtypes for AST types.  For example, consider the syntax-directed
abstract algorithm at the right bottom in Figure~\ref{fig:example}.  The
non-terminal \textit{BindingElement} is the unique token in a alternative of
\textit{FormalParameter} thus we automatically extract the subtype relation:
\textit{BindingElement} $\subtype$ \textit{FormalParameter}.  Based on this
subtype relation, a type check expression $\expr: \ty$ checks whether the
evaluation result of $\expr$ has a type $\ty'$ satisfying $\ty' \subtype \ty$.
We represent semantics of the modified $\ires$ as a state transition system
$(\dom, \trans, \ielemset)$ with the transition relation $\trans \subseteq
\dom \times \dom$.  For presentation brevity, we omit its detail in this
paper and include it in a companion report~\inred{\cite{report}}.
% \[
%   \begin{array}{l@{~}r@{~}c@{~}r@{~}r@{~}l}
%     \text{States}
%     &\elem&\in&\dom&::=&\ctxtset^* \times \labset \times \envset\\
% 
%     \text{Contexts}
%     &\ctxt$\in$\ctxtset&::=&\labset \times \varset \times \envset\\
% 
%     \text{Environments}
%     &\env$\in$\envset&::=&\varset \finmap \valset\\
% 
%     \text{Values}
%     &\val$\in$\valset&::=& \addrset \uplus \addr \uplus \primset\\\
%   \end{array}
% \]

@@

We design a type analysis for the modified $\ires$ based on abstract
interpretation framework~\cite{ai1977, ai1992} with analysis
sensitivity~\cite{sens-toplas}.  We first extend types $\tyset$ to $\etyset$ as
follows:
\[
  \etyset \ni \ety ::=
  \ty \mid
  \func \mid
  \const \mid
  \tnormal(\ety) \mid
  \tabrupt \mid
  \bool \mid
  \str \mid
  \tabsent
\]
There are three different extensions; we add 1) functions and constants
because the original types $\tyset$ do not cover them, 2) normal completions
$\tnormal(\ety)$, abrupt completions $\tabrupt$, Boolean values $\bool$, and
String values $\str$ to increase the precision of type analysis. and 3) the
absent type $\tabsent$ to represent the existence of variables.  Using such
extended types, we define abstract states with flow-sensitivity and
type-sensitivity for parameters as follows:
\[
  \begin{array}{lr@{~}c@{~}c@{~}r@{~}l}
    \text{Abstract States}
    &\aelem&\in&\adom&=&
    \viewset \rightarrow (\aenvset \times \partsof{\viewset})\\

    \text{Views}
    &\view&\in&\viewset&=&
    (\labset \times \etyset^*) \uplus
    (\funcset \times \etyset^*)\\

    \text{Abstract Environments}
    &\aenv&\in&\aenvset&=&\varset \rightarrow \atyset\\

    \text{Abstract Types}
    &\aty&\in&\atyset&=&\partsof{\etyset}\\
  \end{array}
\]
An abstract state $\aelem \in \adom$ is a mapping from views to pairs of
abstract environments and next views.  A view $\view$ is a pair of a label for
the flow-sensitivity and a
sequence of extended types for the type-sensitivity for parameters.  An abstract
environment $\aenv \in \aenvset$ is a mapping from variables to abstract types
and an abstract type is a set of extended types.

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{0.48\textwidth}
    \[
      \begin{array}{r@{~}c@{~}l}
        \asem{\kwlet \; \x = \expr}
        ((\lab, \overline{\ety}), \aenv)(\aelem)
        &=& \todo \\

        \asem{\kwlet \; \x = \kwrl \expr_0 \; \expr_1 \cdots \expr_n \kwrr}
        ((\lab, \overline{\ety}), \aenv)(\aelem)
        &=& \todo \\

        \asem{\kwassert \; \expr}
        ((\lab, \overline{\ety}), \aenv)(\aelem)
        &=& \todo \\

        \asem{\kwif \; \expr \; \lab_0 \; \lab_1}
        ((\lab, \overline{\ety}), \aenv)(\aelem)
        &=& \todo \\

        \asem{\kwreturn \; \expr}
        ((\lab, \overline{\ety}), \aenv)(\aelem)
        &=& \todo \\

        \asem{\refer = \expr}
        ((\lab, \overline{\ety}), \aenv)(\aelem)
        &=& \todo \\
      \end{array}
    \]
    \caption{Instructions with $\asem{\inst}: (\viewset \times
    \aenvset) \rightarrow \adom \rightarrow \adom$}
  \end{subfigure}
  \begin{subfigure}[b]{0.48\textwidth}
    \[
      \begin{array}{r@{~}c@{~}l}
        \tname^? \; \kwcl [\x: \expr]^* \kwcr
        \clist{\expr^*}
        \expr: \ty
        \refer \kwexists
        \expr \bop \expr
        \uop \; \expr
        \refer
        \const
        \prim
      \end{array}
    \]
    \caption{Expressions with $\asem{\expr}: (\viewset \times \aenvset)
    \rightarrow \atyset$}
  \end{subfigure}
  \caption{Abstract semantics of $\ires$ components.}
  \label{fig:example}
\end{figure*}

### 2021-04-12 paper/ase21/overview.tex
[ASE21] A small change in 3.1
[ASE21] temporal dump

\paragraph{Syntax and Semantics} ECMAScript uses a variant of extended BNF
notations to describe JavaScript syntax.  $\jiset$ synthesizes AST structures
and JavaScript parsers from its lexical and syntactic productions.  Our tool
utilizes only AST structures to define AST types and to access corresponding
syntax-directed algorithms.  For JavaScript semantics, ECMAScript contains
abstract algorithms written in a structured natural language.  $\jiset$ compiles
them to functions of $\ires$, an untyped intermediate representation for
ECMAScript.  For brevity, we will formally define syntax and semantics of
a modified $\ires$ in Section~\ref{sec:ires}.

\paragraph{Types} In ECMAScript, types denote not only JavaScript types but also
specification types for ASTs, internal records, and internal lists.  For ASTs,
we use non-terminal names as thier type names and AST types $A$ and $B$ have a
subtype relation $A \subtype B$ when $A$ is an alternative of $B$.  If record
types have names, we treat them as normal types and manually define their fields
according to the latest version of ECMAScript.  For example, fields of the
\code{Realm} record types are defined in ``Table 23: Realm Record Fields'' in
ECMAScript.  Ohterwise, we define structural types for unnamed record types.
For lists, we define \code{Nil} types for empty lists and parametric list types
\code{List[T]} for lists whose elements have \code{T} type.

@@

The type analyzer performs flow-sensitive type analysis with type-sensitivity
for parameters to precisely detect specification bugs.  It consists of two
sub-modules: an \textit{analysis initializer} and an \textit{abstract transfer}.
We use a stack-based worklist algorithm to calculate fixpoint of abstract
transfer and assign the initial set of worklist using analysis initializer.

>>>>

\paragraph{Syntax and Semantics} ECMAScript uses a variant of extended BNF
notations to describe JavaScript syntax.  $\jiset$ synthesizes AST structures
and JavaScript parsers from its lexical and syntactic productions.  Our tool
utilizes only AST structures to define AST types and to access corresponding
syntax-directed algorithms.  For JavaScript semantics, ECMAScript contains
abstract algorithms written in a structured natural language.  $\jiset$ compiles
them to functions of $\ires$, an untyped intermediate representation for
ECMAScript.  We modify $\ires$ to make it more suitable for type analysis, and
formally define its syntax and semantics in Section~\ref{sec:ires}.

\paragraph{Types} In ECMAScript, types denote not only JavaScript types but also
specification types for ASTs, records, and lists.  In abstract algorithms, ASTs
are also values thus they can be stored in the variables or passed as arguments.
For such ASTs, we use their production names as their types and automatically
link corresponding syntax-directed algorithms to them as their fields.
ECMAScript also supports various record types and their fields and possible
values are defined in the corresponding tables.  For example, fields of the
``Realm'' record type are defined in ``Table 23: Realm Record Fields'' in
ECMAScript.  The latests version of ECMAScriptThe contains tens of tables for
record types and they are backward compatible in general.  Thus, we manually
model them to use them in type analysis.  For lists, we define the empty list
type $\clist{}$ and parametric list types $\clist{\tau}$.

@@

The type analyzer performs type analysis with flow-sensitivity and
type-sensitivity for parameters to precisely detect specification bugs.  It
consists of two sub-modules: an \textit{analysis initializer} and an
\textit{abstract transfer}.  We use a stack-based worklist algorithm to
calculate fixpoint of abstract transfer and assign the initial set of worklist
using analysis initializer.
